---
title: "Worksheet 2"
author: "Wendy Nalaza"
date: "2022-10-07"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Worksheet for Programming**

**Instrutions:**

>
+ Use RStudio or the RStudio Cloud accomplish this worksheet. + Save the R script as RWorksheet_lastname#2.R.

>
+ Create your own GitHub repository and push the R script as well as this pdf worksheet to your own repo

Accomplish this worksheet by answering the questions being asked and writing the code
manually

> **Using Vectors**

1. Create a vector using : operator
>a. Sequence from -5 to 5. Write the R code and its output. Describe its output.

```{r x1A}
x <- seq(-5,5)
x
```
b. x <- 1:7. What will be the value of x?

```{r x1B}
x <- 1:7
x
```

2.* Create a vector using seq() function
a. seq(1, 3, by=0.2) # specify step size. Write the R code and its output. Describe the output.

```{r x2A}
seq(1, 3, by=0.2)
```
3.* A factory has a census of its workers. There are 50 workers in total. The following
list shows their ages: 34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18.
```{r x3}
Workers <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
             22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
             24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26, 18)
```
a. Access 3rd element, what is the value?
```{r x3A}
Workers[3]
```
b. Access 2nd and 4th element, what are the values?
```{r x3B}
Workers[2]
Workers[4]
```

6.* The following data shows the diesel fuel purchased by Mr. Cruz.
```{r 4, echo=FALSE}
deisel <- data.frame(
  
  Months = c("Jan", "Feb", "March", "Apr", "May", "June"),
  Php = c(52.50, 57.25, 60.00, 65.00,74.25, 54.00),
  Liter = c(25, 30, 40, 50, 10, 45),
  stringsAsFactors = FALSE)
deisel
```
a. Create a data frame for month, price per liter (php) and purchase-quantity (liter).
Write the codes.
```{r 4A}
deisel$Php
deisel$Liter
```
b. What is the average fuel expenditure of Mr. Cruz from Jan to June? Note: Use weighted.mean(Liter, Php)
```{r 4B}
Php = c(52.50, 57.25, 60.00, 65.00,74.25, 54.00)
Liter = c(25, 30, 40, 50, 10, 45)

weighted.mean(Liter, Php)
```

7.* R has actually lots of built-in datasets. For example, the rivers data “gives the lengths (in miles) of 141 “major” rivers in North America, as compiled by the US Geological Survey”.

a. Type “rivers” in your R console. Create a vector data with 7
elements, containing the number of elements (length) in rivers,
their sum (sum), mean (mean), median (median), variance (var)
standard deviation (sd), minimum (min) and maximum (max).
data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), var(rivers),
sd(rivers), min(rivers), max(rivers))
```{r 5A}
data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), 
          var(rivers),sd(rivers), min(rivers), max(rivers))
data
```
b. What are the results?
```{r 5B}
data
```
c. Write the code and its outputs.
```{r 5C}
data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), 
          var(rivers),sd(rivers), min(rivers), max(rivers))
data
```
8.* The table below gives the 25 most powerful celebrities and their annual pay as ranked by the editions of Forbes magazine and as listed on the Forbes.com website.
Figure 1: Forbes Ranking
```{r 6, echo=FALSE}
 PowerRanking <- 1:25
  Celebrities <- c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods",
          "Steven Spielherg", "Howard Stern", "50 Cent", "Cast of the Sopranos",
          "Dan Brown", "Bruce Springsteen", "Donald Trump", "Muhammad Ali",
          "Paul McCartney","George Lucas","Elton John","David Letterman","Phil Mickelson",
          "J.K Rowling","Bradd Pitt","Peter Jackson","Dr. Phil McGraw",
          "Jay Lenon","Celine Dion","Kobe Bryant")
  Pay <- c(67,90,225,110,90,332,302,41,52,88,55,44,55,40,233,34,40,47,75,25,39,45,32,40,31)
  
  Ranking <- data.frame(PowerRanking, Celebrities, Pay)
  Ranking
```
a. Create vectors according to the above table. Write the codes.
```{r 6A}
 PowerRanking <- 1:25
  Celebrities <- c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods",
          "Steven Spielherg", "Howard Stern", "50 Cent", "Cast of the Sopranos",
          "Dan Brown", "Bruce Springsteen", "Donald Trump", "Muhammad Ali",
          "Paul McCartney","George Lucas","Elton John","David Letterman","Phil Mickelson",
          "J.K Rowling","Bradd Pitt","Peter Jackson","Dr. Phil McGraw",
          "Jay Lenon","Celine Dion","Kobe Bryant")
  Pay <- c(67,90,225,110,90,332,302,41,52,88,55,44,55,40,233,34,40,47,75,25,39,45,32,40,31)
  
  Ranking <- data.frame(PowerRanking, Celebrities, Pay)
  Ranking
```
b. Modify the power ranking and pay of J.K. Rowling. Change power ranking to 15 and pay to 90. Write the codes and its output.
```{r 6B, echo=TRUE}
PowerRanking[19] <- 15;
  PowerRanking  
  
  Pay[19] <- 90;
  Pay

  ArtistRanking <- data.frame(PowerRanking, Celebrities, Pay)
  ArtistRanking  
```
c. Interpret the data.

  **PowerRanking[19] <- 15;**
  **PowerRanking**
  
(Interpret: The first data declares the power ranking at the data ranking of set variables, and the index 19 declares the power ranking of the row, which then changes from 19 to 15.)

**Pay[19] <- 90;**
  **Pay**
  
(Interpret: The second data declares the pay at the data ranking, same as the process at the first data but as the Pay row and declares 19 of the column to change from 75 to 90.)

**ArtistRanking <- data.frame(PowerRanking, Celebrities, Pay)**
  **ArtistRanking**
  
(Interpret: The last data declares another new data as "Artist Ranking" to process the changes inside the data of power ranking, celebrities, and pay.)(Interpret: The second data declares the pay at the data ranking, same as the process at the first data but as the Pay row and declares 19 of the column to change from 75 to 90.)